---
tite: Initial Analyses
author: Melody Royster
date: today
date-format: long
format: typst
editor_options: 
  chunk_output_type: console
---

# Question 1: Where to located your data (a link is preferable), How to download it, and where to save it in order for your code to run?

For my project, I will use two .csv data files from *IPEDS*, the 2023 academic libraries AL2023 and Institutional characteristic, data/hd2022.csv. This files are located within my data file. The original data files can be downloaded from the following links: https://nces.ed.gov/ipeds/datacenter/DataFiles.aspx?sid=5c05e2f4-be4f-4c65-a205-5b539f20afc7&rtid=7

# Question 2: The code

```{r}
library (tidyverse)
library(haven)

data_nolabels <- read_csv("data/hd2022.csv")
data_al <- read_csv("data/al2023.csv")

data_nolabels_al <- left_join(data_nolabels, data_al, by = "UNITID")

data_nolabels_al <- data_nolabels_al %>%
  # mutate(LTCRCLT = log10(LTCRCLT),
  #        LSLIBRN = log10(LSLIBRN)) %>%
  filter(!is.na(LTCRCLT) & !is.na(LSLIBRN))

data_nolabels_al  <- data_nolabels_al %>%
  drop_na(INSTSIZE)

```

I joined my two data files using the `left_join` function and used `UNITID` as the common variable. It is unclear to me how the NA are displaying.

# Question 3: Visualization

```{r}
data_nolabels_al %>%
  ggplot(aes(x =XLTCRCLT)) +
  geom_bar()




```

# 3a

```{r}
data_nolabels_al %>%
  ggplot(aes(x = LTCRCLT,
             y = LSLIBRN)) +
  geom_point(mapping = aes(color = factor ("INSTSIZE")), alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE) +
labs(title = "Academic Library Circulation",
     caption = "Source: IPEDS",
     x = "Library Circulation",
     y = "Librarians") +
  scale_x_continuous(trans='log10') +
  theme_minimal() 

```

This box plot shows the relationship between circulation materials (physical and electronic) and the total number of librarians.

# 3b

```{r}
data_nolabels_al %>%
  ggplot(aes(x = XLTCRCLT, y = XLEXMSTL)) +
  geom_point()


```
